---
title: "Assignment 4"
author: "Carmen Hoyt"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# Load Packages

```{r}
#| message: false
library(sensitivity)
library(tidyverse)
library(gridExtra)
library(purrr)
library(ggpubr)
library(here)
```

# Parameters

1. Windspeeds (v):

- normally distributed
- mean of 300 cm/s
- standard deviation of 50 cm/s

2. Vegetation height (height):

- between 3.5 and 5.5 m (but any value in
that range is equally likely)

3. k_d and k_0 parameters: 

- normally distributed
- standard deviation of 1% of their default values

# Analysis 

a. Use the Sobel approach to generate parameter values
for the 4 parameters

```{r}
#| message: false
source(here("Catm.R"))

# generate two examples of random number from parameter distributions

np <- 1000
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.01, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.01, n = np)
# in meters/s
v <- rnorm(mean = 3, sd = 0.5, n = np)
height <- runif(min = 3.5, max = 5.5, n = np)

X1 <- cbind.data.frame(k_o, k_d, v, height = height)

# repeat sampling
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.01, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.01, n = np)
# in meters/s
v <- rnorm(mean = 3, sd = 0.5, n = np)
height <- runif(min = 3.5, max = 5.5, n = np)

X2 <- cbind.data.frame(k_o, k_d, v, height = height)

# there are different versions of sobol functions that have different approaches for estimating parameters and indices, we use an approach implemented by jansen

sens_Catm_Sobol <- sobolSalt(model = NULL, X1, X2, nboot = 100)

#str(sens_Catm_Sobol)
# Take a look at the Sobol generated sensitivity object
# your parameters sets for sensitivity analysis are in X

#sens_Catm_Sobol$X
```

b. Run the atmospheric conductance model for these
parameters

```{r}
# run model for all parameter sets
# make sure you give the parameters names

parms <- as.data.frame(sens_Catm_Sobol$X) # turn results into df
colnames(parms) <- colnames(X1) # give it original column names
res <- pmap_dbl(parms, Catm) # pmap parameters to the function (run model)

# length(res) == nrow(X)

sens_Catm_Sobol <- sensitivity::tell(sens_Catm_Sobol, res, res.names = "ga") 
```

c. Plot conductance estimates in a way that accounts for
parameter uncertainty

```{r}
# graph two most sensitive parameters
both <- cbind.data.frame(parms, gs = sens_Catm_Sobol$y) #gs is atmostpheric conductance (results)

# look at overall gs sensitvity to uncertainty
ggplot(both, aes(x = gs)) +
  geom_histogram() +
  geom_vline(xintercept = mean(both$gs), col = "cyan")

# use second most sensitive parameter (using most important as color)
ggplot(both, aes(k_o, gs, col = k_d)) +
  geom_point() +
  labs(y = "Conductance (mm/s)", x = "k_d parameter")
```

d. Plot conductance estimates against windspeed use the
parameter that is 2nd in terms of total effect on
response

```{r}
# look at response of conductance to the two interesting variables
ggplot(both, aes(v, gs, col = height)) +
  geom_point() +
  # conductance as a function of windspeed (height isnt making much diff i.e. color)
  labs(y = "Conductance (mm/s)", x = "Windspeed")
```

e. Estimate the Sobel Indices for your output

```{r}
# quanitify Sobol indices
# tell needs Sobel object (sens_Catm_Sobol) and results (res)
sens_Catm_Sobol <- sensitivity::tell(sens_Catm_Sobol, res, res.names = "ga") 
parms <- as.data.frame(sens_Catm_Sobol$X)
colnames(parms) <- colnames(X1)
res <- pmap_dbl(parms, Catm)

sens_Catm_Sobol <- sensitivity::tell(sens_Catm_Sobol, res, res.names = "ga")

# main effect:  partitions variance (main effect without co-variance) - sums approximately to one
sens_Catm_Sobol$S
# useful to add names
row.names(sens_Catm_Sobol$S) <- colnames(parms)
sens_Catm_Sobol$S # S is the main effect

# total effect - accounts for parameter interactions
row.names(sens_Catm_Sobol$T) <- colnames(parms)
sens_Catm_Sobol$T # T is the total effect

# Both the main effect and total effect can tell us something about how the parameter influences results

print(sens_Catm_Sobol)
```


f. Comment on what this tells you about how
atmospheric conductance and its sensitivity to
variation in windspped differs in this setting as
compared to the setting that we examined in class
where windspeed was lower and less variable and
vegetation was taller.


